---
title: 'Tutoring CIND123 Winter 2018 - Assignment #2'
author: "Faizan Khalid Mohsin; 997157570"
date: "February 8, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use RStudio for this assignment.  Edit the file `assignment-2.Rmd` and insert your R code where wherever you see the string "INSERT YOUR ANSWER HERE" 

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.  
 
When your are done with your answers and before submitting, save the file with the following naming convention :your **Lastname_firstname** 
 
Submit **both** the rmd and the pdf output(or word or html) files, failing to submit **both** will be subject to mark deduction. 
 
This assignment may make use of data provided by the `ISwR` package. 
 
```{r} #library(ISwR) ``` 
 
## Sample Question and Solution 
 
Use `seq()` to create the vector $(1,2,3,\ldots,10)$. 
 
```{r} seq(1,10) ``` 
 
****** \newpage ## Question 1 (50%) 
 
Consider the probability distribution associated with rolling 3 fair dice. We can label the faces of a single die using the numbers from 1 to 6. We can therefore label the simple events in this distribution by triples of numbers from 1 to 6. Let `d1`, `d2`, and `d3` represent the labels on each of the dice. 
 
a) Set `d1` to the sequence (1,2,...,6) repeated 36 times. 
 
```{r} d1 = rep(1:6, 36) ``` 
 
b) Set `d2` to the sequence consisting of 6 repetitions of the sequence in which each of the numbers (1,2,...,6) is repeated 6 times. 
 
```{r} d2 = rep(rep(1:6, each = 6), 6) ``` 
 
c) Set `d3` to the sequence in which each of the numbers (1,2,...,6) is repeated 36 times. 
 
```{r} d3 = rep(1:6, each = 36) ``` d) Create a new data frame `three.dice` from `d1`, `d2`, and `d3` and print it. Visually confirm that there are $6 \times 6 \times 6 = 216$ rows and each row contains a unique combination of dice labels. 
 
```{r} three.dice= cbind(d1, d2, d3) ;print(three.dice) ``` e) Since the dice are fair and independent, each simple event has the same probability, namely $\frac{1}{216}$. Add the column `P` to the data frame with this value. ```{r} p = rep(1/216, 216) ; three.dice = cbind(d1, d2, d3, p)``` f) Add a new column `sum` equal to the sum of the dice labels ```{r} sum = d1+d2+d3```. Add another new column `mean` equal to the average of the dice labels ```{r} mean = (d1+d2+d3)/3 ```. 

Final Three.Dice data frame is:
``` three.dice = cbind(d1, d2, d3,p, sum, mean)
 
```{r} hist(three.dice$sum) ``` g) Plot a probability histogram of `three.dice$sum`. 
 
```{r} #Insert your answer here ``` h) Compute the probability that the sum of the dice is greater then 12 and less than 18. 
 
**HINT**: Use `subset()` to select the events and sum `P`. 
 
```{r} #Insert your answer here ``` i) Compute the probability that the sum is even. 
 
```{r} #Insert your answer here 
``` j) Compute the probability that the mean is exactly 4. 
 
```{r} #Insert your answer here ``` ****** \newpage ## Question 2 (50%) a)You have two groups of distinctly different items, 10 in the first group and 8 in the second. If you select one item from each group, how many different pairs can you form? ```{r} #use the mn rule #Insert your answer here ``` 
 
b)Evaluate the following permutation $P_{3}^5$ ```{r} #Insert your answer here ``` 
 
c)Evaluate the following combinations $C_{3}^5 + C_{2}^5$ ```{r} #Insert your answer here ``` d)In how many ways can you select five people from a group of eight if the order of selection is important? ```{r} #Since order is important, you might need to use permutations #Insert your answer here ``` e)In how many ways can you select two people from a group of 20 if the order of selection is not important? ```{r} #Since order is unimportant, you might need to use combinations #Insert your answer here ``` 
 
 
END of Assignment #2. 